---
title: "Some gun violence analysis with Wikipedia data"
output: html_document
---

This analysis is a reperformance of http://nbviewer.ipython.org/gist/minrk/4358066

The result expected is to analyze the following relationship ratios:
- Homicides using guns
- Intentional homicides
- Countries by per-capita GDP

To reperform the analysis, was considered the following data:

- totals.csv - List of countries by intentional homicide rate - https://en.wikipedia.org/wiki/List_of_countries_by_intentional_homicide_rate
- guns.csv - List of countries by firearm-related death rate - https://en.wikipedia.org/wiki/List_of_countries_by_firearm-related_death_rate
- gdp.csv - Countries by GDP per capita - https://en.wikipedia.org/wiki/List_of_countries_by_GDP_(PPP)_per_capita

```{r, echo=FALSE}
library(data.table)
setwd('~/Desktop/gdp/')
```

Loading and cleaning the data:

```{r}
totals = data.table(read.csv('totals.csv', sep = ';', header = TRUE))
setnames(totals, c("Country","Intentional_Homicides_Rate","Intentional_Homicides_Total_Count","Intentional_Region","Intentional_Subregion","Intentional_Homicides_Year"))

guns = data.table(read.csv('guns.csv', sep = ';', header = TRUE)[,c(1,2,4,5,12)])
setnames(guns, c("Country","Homicides_Total","Gun_Homicides_Total","Gun_Homicides_Year","Gun_Homicides_Sources_and_Notes"))

gdp = data.table(read.csv('gdp.csv', sep = ';', header = TRUE))
setnames(gdp, c("GDP_Rank","Country","GDP_Int.dolar","GDP_Year"))

gdp$GDP_Rank = as.numeric(gdp$GDP_Rank)
```

Merging the datasets

```{r}
data = merge(totals, guns, by="Country")
data = merge(data, gdp, by="Country")
```

Extracting the percenage of homicides using guns from the total of homicides

```{r}
data$Gun_Percent = 100 * data$Gun_Homicides_Total / data$Homicides_Total
```

Sorting by "Gun_Homicides_Total" and presenting the top 15 countries by total of homicides related to guns

```{r}
data = data[order(Gun_Homicides_Total, decreasing = TRUE)]
head(data[,c(1,7,8,10), with = FALSE], 15)
```

Sorting by GDP ranking and presenting the top 30 countries by total of homicides related to guns
```{r}
top = data[order(GDP_Rank)][1:30,c(1,7,8), with = FALSE]
top = top[order(Gun_Homicides_Total, decreasing = TRUE)]
head(top, 10)

#df = head(top, 10)[,1:2, with = F]

#barplot(df, 
#        main="Countries by Guns Homicides and Total Homicides",
#        xlab="Countries",
#        ylab ="per 100.000 people",
#        col=c("darkblue","red"),
#        legend = rownames(df))
```

Adding USA, Canada, and Mexico to all of Europe

```{r}
americas_countries = c("United States", "Canada", "Mexico")
reg_data = data[Intentional_Region %in% "Europe"]
reg_data = rbind(reg_data, data[Country %in% americas_countries])

top_reg = reg_data[order(GDP_Rank)][1:30,c(1,7,8), with = FALSE]
top_reg = top_reg[order(Gun_Homicides_Total, decreasing = TRUE)]
head(top_reg, 25)
```

Comparing US, Canada and UK

```{r}
americas_countries = c("United States", "Canada", "United Kingdom")
reg2_data = data[Country %in% americas_countries]

top_reg2 = reg2_data[order(GDP_Rank)][,c(1,7,8), with = FALSE]
top_reg2 = top_reg2[order(Gun_Homicides_Total, decreasing = TRUE)]
top_reg2
```
Normalize to the US numbers (inverse)

```{r}
americas_countries = c("United States", "Canada", "United Kingdom")
reg2_data = data[Country %in% americas_countries]

top_reg2 = reg2_data[order(GDP_Rank)][,c(1,7,8), with = FALSE]
top_reg2 = top_reg2[order(Gun_Homicides_Total, decreasing = TRUE)]
top_reg2
```